---
title: "Price predictors of used cars"
author: "Bilal Naseem, Franco Doi (30270314), Gautham Chandra Shekariah, Joao Bertti Targino, Romith Bondada"
date: "2025-03-30"
output: 
  bookdown::word_document2:
    toc: true
    fig_caption: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: sentence
header-includes:
  - \usepackage{amsmath}
  - \usepackage{hyperref}
  - \usepackage{graphicx}
  - \usepackage{float} 
  - \usepackage{listings}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \setcounter{figure}{0}
  - \lstset{breaklines=true, xleftmargin=20pt, basicstyle=\tiny, frame=single}
geometry: margin=1in
fontsize: 11pt
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(mosaic)
library(tidyverse)
library(webshot2)
library(latex2exp)
library(olsrr)
library(bookdown)


library(rmarkdown)
library(officer)

library(corrplot)
library(ggplot2)
library(gridExtra)
library(scales)
library(lubridate)
library(knitr)
library(reshape2)

```

\newpage
**DATA 603 (W25)**  
**Final Report**

**Title**: Price Predictors of Used Cars  

**Group Members**:  

*Full names and UCIDS*

- Bilal Naseem  
- Franco Doi (30270314)  
- Gautham Chandra Shekariah  
- Joao Bertti Targ

\newpage
# INTRODUCTION 

## Motivation

### Context
*Clearly state the context and applied domain(s) of your project.*

The automotive market for pre-owned or used vehicles is dynamic and often unpredictable. As noted in the study "Predicting Car Prices: A Comparative Analysis of Machine Learning Techniques" (Abdulhakeem et al., 2023), the accurate prediction of car prices is crucial for various stakeholders in the automotive industry, including manufacturers, dealers, and consumers.

The used car market presents unique challenges for price prediction due to the numerous factors that influence a vehicle's value beyond just its base specifications. These factors include age, mileage, condition, brand reputation, market demand, and regional economic conditions. Traditional methods of car valuation often fail to capture the complex relationships between these variables.

Machine learning approaches offer a promising solution to this problem by analyzing patterns in historical data to generate accurate price predictions. The referenced study compares several machine learning algorithms including Linear Regression, Random Forest, and XGBoost to determine which provides the most accurate predictions for used car prices.


### Problem
*What problem are we going to address and propose a data/visual analytics solution to?*

This project aims to develop a predictive model that estimates the sale price of a pre-owned vehicle based on a variety of numerical and categorical features. The motivation stems from a personal and practical interest: as individuals who may buy or sell used cars, we want to ensure we are not overpaying or underselling. A reliable prediction model could help users make more informed financial decisions in the used car market. 

### Challenges
*Why is the problem hard? Provide examples*

Predicting the price of a used vehicle involves several challenges. First, the data may contain missing values, inconsistent formats, or outliers that need to be cleaned and handled appropriately. Second, the relationship between features (such as brand, age, mileage, and condition) and price are often nonlinear and complex. Third, combining both categorical and numerical variables in a model requires careful preprocessing and feature engineering. Finally, ensuring the model generalizes well to unseen data demands rigorous validation and tuning.

## Objectives

### Overview
*What is the overall intent of the project?*

The objective of this project is to build a predictive model that can accurately estimate the sale price of a pre-owned/used car. This model will leverage a combination of numerical and categorical variables to understand the factors that influence pricing in the used car market. By applying data analysis and machine learning techniques, we aim to create a practical tool for evaluating whether a vehicle is fairly priced—either for personal use or broader market insight.

### Goals & Research Questions
*Clearly state each of your data/visual analytics goals and research questions.*

#### Goals

1. To analyze the key variables that affect the sale price of used cars.
2. To preprocess and explore the dataset for patterns, correlations, and insights.
3. To build, evaluate, and compare predictive models that estimate car prices based on input features.
4. To create a tool that can help users make more informed decisions when buying or selling a used vehicle.

#### Research Questions

1. What are the most influential features in determining a used car’s sale price?
2. How accurately can we predict the price using machine learning algorithms?
3. Are there noticeable trends or patterns in pricing based on make, model, age, or mileage?
4. Can the model help identify potentially overpriced or underpriced vehicles?













\newpage
# METHODOLOGY  

## Data
*Describe the data set(s) related to your project, source(s), how data was collected, number of attributes and a brief description of each attribute.*

### Data Source
The dataset used in this project is the "Used Car Price Prediction Dataset" obtained from Kaggle (https://www.kaggle.com/datasets/taeefnajib/used-car-price-prediction-dataset). This dataset contains information about used cars listed for sale, including various attributes that potentially influence the selling price.

### Data Collection

The dataset is available online and is licensed under Attribution 4.0 International, which allows us to use for academic purposes. 

The data was likely collected from online car listings or dealership inventory, though the specific collection methodology isn't explicitly stated in the Kaggle source. Without additional metadata, it's difficult to determine if the sampling was random or if there are geographic or temporal biases in the collection.

### Dataset Size
The dataset consists of 6,019 observations (rows), each representing a specific used car listing. It contains 13 variables (columns) capturing various car attributes that might influence pricing.

### Attributes Description
Key features in the dataset include:

- Vehicle characteristics (make, model, year, mileage)
- Physical attributes (engine capacity, fuel type, transmission type)
- Market factors (listing price)

Detailed explanation of each attribute is below:

-	**Price**: The listed selling price of the used car (in currency units). This is the target variable we aim to predict.
-	**Brand**: The manufacturer or make of the car (e.g., Toyota, Honda, Ford, BMW, Mercedes-Benz). This is a categorical variable representing different car manufacturers.
-	**Model**: The specific model of the car offered by the manufacturer. This is a categorical variable with numerous unique values depending on the brand.
-	**Year**: The manufacturing year of the vehicle. This numerical variable indicates the car's age, which typically has a strong correlation with its price.
-	**Mileage**: The total distance the car has been driven in miles. This numerical variable is a key indicator of vehicle wear and significantly impacts resale value.
-	**Fuel Type**: The type of fuel the car uses (e.g., Gasoline, Diesel, Electric, Hybrid). This categorical variable can significantly influence pricing due to differences in fuel economy and operational costs.
-	**Engine Capacity**: The engine’s specifications including size, technology, number of cylinders, generally correlates with the car's power and fuel consumption.
-	**Transmission Type**: The type of transmission system in the car (e.g., Automatic, Manual, CVT, Semi-automatic). This categorical variable affects both driving experience and maintenance costs.
- **Exterior Color**: Exterior color of the car. The exterior color significantly influences a vehicle's marketability and resale value.
- **Interior Color**: Interior color of the car. Interior color affects both buyer perception and practical considerations.
- **Accident**: Accident history of the car. A vehicle's accident history can dramatically impact its value and desirability.

The dataset provides a comprehensive representation of the used car market with attributes covering the most significant factors that influence car pricing. These variables include both technical specifications (engine capacity, mileage) and market segmentation factors (brand, model), allowing for a thorough analysis of price determinants.

## Approach
*What approach are you using for data/visual analytics solution? Why do you think it will work well?*

For our used car price prediction project, we are implementing a comprehensive data and visual analytics solution that combines exploratory data analysis (EDA), statistical modeling, and rigorous validation techniques. Our approach consists of several integrated components:


### Exploratory Data Analysis with Visualizations

We begin with an extensive EDA that uses a variety of visualizations to understand the dataset's characteristics:

a)	**Distribution analyses** of continuous variables (price, mileage, year) using histograms and density plots 
b)	**Relationship visualizations** through scatter plots, box plots, and correlation matrices 
c)	**Category-based comparisons** using bar charts and grouped box plots to assess how categorical variables like brand, model, and fuel type affect prices
d)	**Multi-dimensional analyses** including faceted plots and color-coded visualizations to reveal complex interactions between variables


### Modeling with Multiple Linear Regression

For prediction, we employ a multiple linear regression model with several enhancements:

a)	**Interaction terms** between significant variables to account for conditional effects 
b)	**Polynomial transformations** of key numerical predictors to capture non-linear relationships 
c)	**Feature engineering** including derived variables like car age and miles-per-year 
d)	**Comprehensive model selection** using techniques like best subset selection, adjusted R-squared, AIC, and BIC criteria


### Statistical Validation and Diagnostics

We implement a rigorous validation framework:

a)	**Assumption testing** for linearity, independence, homoscedasticity, and normality 
b)	**Multicollinearity detection** using Variance Inflation Factors (VIF) 
c)	**Influential observation analysis** using Cook's distance and DFBETAS 
d)	**Cross-validation** with train/test splits to confirm model generalizability 
e)	**Performance metrics** including RMSE and R-squared

### Results Visualization and Interpretation
Our final analytics layer includes:

a)	**Coefficient visualization** to show the impact of each predictor 
b)	**Actual vs. predicted** plots to illustrate model accuracy 
c)	**Residual analysis visualizations** to identify potential improvements 
d)	**Interactive visualizations** that allow for dynamic exploration of model predictions

### Why this works
Our integrated approach is particularly well-suited for the used car price prediction problem for several reasons:

1)	**Addresses Domain-Specific Challenges**

    The used car market has well-known non-linear aspects (e.g., depreciation curves) and complex interactions (e.g., how premium brands depreciate differently) that our approach specifically targets through polynomial terms and interaction effects.
  
2)	**Balances Complexity and Interpretability**
  
    While machine learning algorithms like random forests might achieve high accuracy, linear regression with carefully selected enhancements provides comparable performance while maintaining interpretability—a crucial factor when the goal is not just prediction but understanding what drives car prices.
  
3)	**Provides Rigorous Statistical Foundation**
  
    Our statistical validation ensures that the model meets all necessary assumptions, preventing misleading results and ensuring the reliability of predictions across different market segments.
  
4)	**Leverages Visual Analytics for Insight Generation**
  
    By integrating visualization throughout the analysis pipeline—from initial exploration to final model evaluation—we facilitate deeper insights about the used car market that purely algorithmic approaches might miss.
  
5)	**Enables Practical Application**

    The final model delivers actionable insights for various stakeholders:

    +	Buyers can understand fair market values based on car attributes
    +	Sellers can set competitive yet profitable prices
    +	Dealerships can optimize inventory pricing strategies
    +	Market analysts can identify trends and anomalies in the used car market

By combining robust statistical modeling with comprehensive visual analytics, our approach not only predicts prices with high accuracy but also provides valuable insights into the factors that drive the used car market.

## Workflow
*What steps (workflow task list) are required? Which of these steps is particularly hard? What to do if the hard steps don't work out*

The workflow for building a robust used car price prediction model is as follows:

1. **Exploratory Data Analysis (EDA)**

    The first step involves understanding the structure, distribution, and relationships within the dataset. This includes identifying missing values, outliers, data types, and potential data quality issues. EDA also helps form initial hypotheses and informs feature engineering strategies.
    
2. **Data Cleaning and Preprocessing**

    This includes handling missing values, correcting inconsistent entries, and converting categorical features to a usable format (e.g., one-hot encoding or label encoding). Normalization or scaling of numerical features may also be applied here.

3. **Feature Selection and Collinearity Check**

    Highly correlated features can lead to multicollinearity, which may negatively impact some models. Identifying and removing redundant features helps improve model performance and interpretability.

4. **Modeling**

    Build a linear regression model. Include polynomial terms or interactions if they improve performance. 

5. **Evaluation**

    Use metrics such as MAE, RMSE, and R² on validation data to assess model accuracy. Validate with a train/test split or cross-validation.

6. **Iterative Tuning**

    Based on validation results, revisit earlier steps—adjust feature engineering, try new models, or fine-tune hyperparameters. This iterative loop continues until a satisfactory model is achieved.
    
As industry standard value $\alpha$ will be set as $0.05$ throughout the project.

## Challenges and Potential issues

Given the nature of used car listings, potential sampling biases might include:

* Over-representation of popular car models
* Limited geographic coverage if data was collected from specific regions
* Possible exclusion of private sales if only dealership data was used

The dataset appears designed primarily for developing machine learning models to predict used car prices based on vehicle attributes.


## Contributions

*Briefly describe the group members’ workload distribution and responsibilities.*















\newpage
# MAIN RESULTS OF THE ANALYSIS

## Results

*What do my results indicate? Do you have any unexpected results? Please elaborate.*

### Data cleaning & transformation

The numerical features price and milage were string like “\$10,300” for price and “5,1000 mi” for milage.
The characters “\$” and “,” were removed for price and “,”, “ “ and “mi” were removed for milage prior to converting them to numerical features.
Column milage was renamed to mileage for easier understanding.


```{r}
# 1. Data Loading and Initial Exploration
# ---------------------------------------

zip_url <- "https://www.kaggle.com/api/v1/datasets/download/taeefnajib/used-car-price-prediction-dataset"
# Create a temporary file to download the ZIP
temp_file <- tempfile(fileext = ".zip")

# Download the file
download.file(zip_url, temp_file, mode = "wb")

# List files inside the ZIP
csv_file_name <- unzip(temp_file, list = TRUE)$Name[1]  # choose appropriate file if more than one

used_cars_data <- read.csv(unz(temp_file, csv_file_name))
```

```{r}
# remove string and convert string to number
used_cars_data$price <- gsub("[$,]", "", used_cars_data$price)
used_cars_data$price <- as.integer(used_cars_data$price)

# remove string and convert string to number
used_cars_data$milage <- gsub("[, mi]", "", used_cars_data$milage)
used_cars_data$milage <- as.integer(used_cars_data$milage)

#renaming column
names(used_cars_data)[names(used_cars_data) == "milage"] <- "mileage"

# Replace missing 'fuel_type' values with "electric"
used_cars_data$fuel_type[is.na(used_cars_data$fuel_type)] <- "electric"
```

Categorical attributes are converted into type factor to avoid wrong interpratation by R
```{r}
used_cars_data$brand <- as.factor(used_cars_data$brand)
used_cars_data$model <- as.factor(used_cars_data$model)
used_cars_data$fuel_type <- as.factor(used_cars_data$fuel_type)
used_cars_data$engine <- as.factor(used_cars_data$engine)
used_cars_data$transmission <- as.factor(used_cars_data$transmission)
used_cars_data$ext_col <- as.factor(used_cars_data$ext_col)
used_cars_data$int_col <- as.factor(used_cars_data$int_col)
used_cars_data$accident <- as.factor(used_cars_data$accident)
used_cars_data$clean_title <- as.factor(used_cars_data$clean_title)
```

### EDA
#### Summary
After data cleaning the dataset has 4009 rows and 12 attributes, and corresponding data types.
```{r}
# Display basic information about the dataset
str(used_cars_data)
```

Below is the summary of the dataset
```{r}
summary(used_cars_data)
```

There is no missing values nor duplication.
```{r}
# Check for missing values
missing_values <- colSums(is.na(used_cars_data))
cat("\nMissing values per column:\n")
print(missing_values[missing_values > 0])
```
```{r}
# Check for duplicates
cat("\nNumber of duplicate rows:", sum(duplicated(used_cars_data)), "\n")
```

#### Univariate Analysis
##### Price distribution

```{r}
figure_no <- 1
ggplot(used_cars_data, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Distribution of Car Prices",
       x = "Price ($)",
       y = "Count",
       caption = paste("Figure", figure_no,": Histogram showing the distribution of used car prices with 50 bins.")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5, face = "italic"))
```

The plot above shows that most car prices are less than 1,000,000, and clearly have outliers, which makes the analysis of distribution hard to visualize.

To visualize the distribution with less impact of the outliers, the car prices are transformed by log:

```{r}
figure_no <- figure_no + 1
ggplot(used_cars_data, aes(x = log(price))) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  labs(title = "Distribution of Log-Transformed Car Prices",
       x = "Log(Price)",
       y = "Count",
       caption = paste("Figure", figure_no, ": Log transformation applied to price values to better visualize the distribution pattern.")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5, face = "italic"))
```

The figure above shows that the distribution is slightly skewed to the right, probably caused by the outliers.

##### Manufacturing year distribution

```{r}
figure_no <- figure_no + 1
ggplot(used_cars_data, aes(x = model_year)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Distribution of Car's manufacturing year",
       x = "Manufacturing year",
       y = "Count",
       caption = paste("Figure",figure_no,": Histogram showing the distribution of used car manufacturing year with 50 bins.")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5, face = "italic"))
```


The figure above shows distribution for the model year on the left and the mileage on the right.

The model year is skewed to the left, where most of the cars are newer.

##### Mileage distribution
```{r}
figure_no <- figure_no + 1
ggplot(used_cars_data, aes(x = mileage)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Distribution of Car's mileage",
       x = "Mileage",
       y = "Count", 
       caption = paste("Figure",figure_no,": Histogram showing the distribution of used car mileage with 50 bins.")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5, face = "italic"))
```


The Mileage is skewed to the right, where most of the cars were driven less.

##### Brand type Distribution

```{r}
# Function to create bar plots for categorical variables
plot_categorical_var <- function(data, var_name, figure_no, max_categories = 25) {
  # Increment figure number
  figure_no <- figure_no + 1
  
  # Get top N categories if there are too many
  if(length(unique(data[[var_name]])) > max_categories) {
    top_cats <- names(sort(table(data[[var_name]]), decreasing = TRUE)[1:max_categories])
    plot_data <- data %>%
      mutate(temp_var = ifelse(.data[[var_name]] %in% top_cats, 
                              as.character(.data[[var_name]]), 
                              "Other")) %>%
      count(temp_var) %>%
      rename(!!var_name := temp_var)
    
    # Create caption for plots with "Other" category
    caption_text <- paste("Figure", figure_no, ": Showing top", max_categories, 
                         "categories of ",var_name,". All remaining categories grouped as 'Other'.")
  } else {
    plot_data <- data %>% 
      count(.data[[var_name]])
    
    # Create caption for plots with all categories shown
    caption_text <- paste("Figure", figure_no, ": Distribution of all", 
                         length(unique(data[[var_name]])), "categories.")
  }
  
  # Create the plot using tidy evaluation
  p <- ggplot(plot_data, aes(x = .data[[var_name]], y = n)) +
    geom_bar(stat = "identity", fill = "seagreen", alpha = 0.7) +
    labs(title = paste("Count of Cars by", var_name),
         x = var_name,
         y = "Count",
         caption = caption_text) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 45, hjust = 1),
          plot.caption = element_text(hjust = 0.5, face = "italic"))
  
  # Return both the plot and the updated figure number
  return(list(plot = p, figure_no = figure_no))
}
```


```{r}
result <- plot_categorical_var(used_cars_data, "brand", figure_no)
result$plot
```

Visually the distribution of brands is not uniform, which indicates unbalanced data regarding brand and may bias the regression.

Brands has no invalid or missing value.

Since every car has a brand and a model, therefore brand and model are highly correlated and will violate the multicollinearity assumption. Therefore brand will be removed since has less granularity than model.


##### Model distribution
```{r}
result <- plot_categorical_var(used_cars_data, "model", figure_no)
result$plot
```

Visually the distribution of car models looks uniform, with count of other models being the highest.

The models has no invalid value.

##### Fuel type distribution
```{r}
result <- plot_categorical_var(used_cars_data, "fuel_type", figure_no)
result$plot
```

Gasoline is the most popular fuel type, and there are two unexpected fuel type:
1. Dash (-) : it is an error entry or data was unavailable, in total are 45 rows.
2. Empty string: it is for electric cars, in total are 170 rows.
3. "not supported": it is for electric cars, in total are 2 rows.

For the electric cars, the fuel type is set to "Electric".


For the rows with "Dash", it is observed that engine is also a "Dash", making it impossible to sanitize and will be left for now.


##### Engine type distribution
```{r}
result <- plot_categorical_var(used_cars_data, "engine", figure_no)
result$plot
```

Visually the distribution of car engine looks uniform, with count of other engine types being the highest.

The engine attribute has '-' which are missing values, on the same rows as missing values for fuel type.

This attribute is very descriptive and does not have a clear standard like '3.0 Liter' and '3.0L' and might be hard to be used for the final model.


##### Transmission type distribution
```{r}
result <- plot_categorical_var(used_cars_data, "transmission", figure_no)
result$plot
```

Visually the distribution of car transmission are mostly A/T, which seems to be short of Automatic Transmission.

The engine attribute has '-' which are missing values for 4 rows.

This attribute is very descriptive and does not have a clear standard like '2-Speed A/T' and '2-Speed Automatic' and might be hard to be used for the final model.

##### Exterior color distribution
```{r}
result <- plot_categorical_var(used_cars_data, "ext_col", figure_no)
result$plot
```

Visually the distribution of car exterior color have few colors that are ver common.

The engine attribute has '-' which are missing values for 15 rows. This information would be easily obtained by inspecting the vehicle and updating the data but is not possible in this dataset.

This attribute has typo like 'Blu', and is very descriptive and does not have a clear standard like 'Blue' and 'BLUE', or 'C / C' and might be hard to be used for the final model.

##### Interior color distribution
```{r}
result <- plot_categorical_var(used_cars_data, "int_col", figure_no)
result$plot
```

Visually the distribution of car interior color have few colors that are ver common.

The engine attribute has '-' which are missing values for 133 rows. This information would be easily obtained by inspecting the vehicle and updating the data but is not possible in this dataset.

This attribute has typo like 'Blk', and is very descriptive and does not have a clear standard like 'Black' and 'BLACK', or 'Sport' and might be hard to be used for the final model.


##### Accident history distribution
```{r}
result <- plot_categorical_var(used_cars_data, "accident", figure_no)
result$plot
```

Visually the distribution of car accident history have no accident reported.

There are 133 missing rows with missing value. Since is not trivial to get a car's accident history, there is a chance that these rows will be dropped or the column will be dropped. There is high chance that this column is higly correlated with accident attribute.

##### Clean title distribution
```{r}
result <- plot_categorical_var(used_cars_data, "clean_title", figure_no)
result$plot
```

Visually the distribution of car's clean title have "Yes" as clean title reported.

There are 596 missing rows with missing value. 

Since a car only has clean title if has no accident, clean_title and accident are highly and will violated multicollinearity assumption. Since clean_title has more missing values than accident, this column will be dropped.


```{r}
pairs(~price+model_year+mileage, data=used_cars_data, panel = panel.smooth)
```

```{r}
# Example: Check how horsepower varies by model
boxplot(mileage ~ brand, data = used_cars_data, las = 2,
        main = "Horsepower by Model", ylab = "Horsepower", xlab = "Model", col = "orange")
```

```{r}
library(ggplot2)

ggplot(used_cars_data, aes(x = brand, y = mileage)) +
  geom_violin(fill = "lightblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Horsepower by Model", y = "Horsepower", x = "Model")
```

```{r}
anova_model <- aov(mileage ~ brand, data = used_cars_data)
summary(anova_model)
```



```{r}
library(car)  # More robust VIF calculation
library(lmtest)  # For additional regression diagnostics
library(olsrr)  # Additional regression diagnostics

firstordermodel <- lm(price ~ model_year + mileage + brand + model + 
                      fuel_type + engine + transmission + 
                      ext_col + int_col + accident + clean_title, 
                      data = used_cars_data)
```


```{r}
options(max.print = 1000000000)
writeLines(capture.output(alias(firstordermodel)), "alias_output.txt")
```

```{r}
df <- used_cars_data
# Calculate the frequency of each brand
brand_freq <- table(df$model)

# Sort by frequency and keep the top 3 brands, others will be "Other"
top_brands <- names(sort(brand_freq, decreasing = TRUE)[1:10])
df$used_cars_data <- ifelse(df$model %in% top_brands, df$model, 'Other')

# Check the updated dataframe
table(df$used_cars_data)
```

```{r}
df <- used_cars_data
# Create dummy variables for 'brand'
df_dummies <- model.matrix(~ model - 1, data = df)

# Perform PCA
pca_result <- prcomp(df_dummies)
summary(pca_result)

# Extract the proportion of variance explained by each component
explained_variance <- pca_result$sdev^2 / sum(pca_result$sdev^2)

# Plot the cumulative explained variance
cumulative_variance <- cumsum(explained_variance)

# Plot the cumulative variance
plot(cumulative_variance, type = "b", xlab = "Number of Components", ylab = "Cumulative Explained Variance", main = "Cumulative Explained Variance vs. Number of Components")

# Programmatically select the number of components that explain 90% of the variance
threshold <- 0.90
n_components <- which(cumulative_variance >= threshold)[1]

# Show the number of components selected
cat("Number of components explaining", threshold * 100, "% of variance:", n_components, "\n")

# Use only the selected number of components in the model
selected_components <- pca_result$x[, 1:n_components]
```

```{r}
str(used_cars_data)
```



```{r}
# Convert categorical variables ('model' and 'ext_color') to dummy variables
df_dummies_brand <- model.matrix(~ brand - 1, data = used_cars_data)  # Dummy variables for 'model'
df_dummies_model <- model.matrix(~ model - 1, data = used_cars_data)  # Dummy variables for 'model'
df_dummies_fuel_type <- model.matrix(~ fuel_type - 1, data = used_cars_data)  # Dummy variables for 'model'
df_dummies_engine <- model.matrix(~ engine - 1, data = used_cars_data)  # Dummy variables for 'model'
df_dummies_transmission <- model.matrix(~ transmission - 1, data = used_cars_data)  # Dummy variables for 'model'
df_dummies_ext_color <- model.matrix(~ ext_col - 1, data = used_cars_data)  # Dummy variables for 'ext_color'
df_dummies_int_color <- model.matrix(~ int_col - 1, data = used_cars_data)  # Dummy variables for 'ext_color'
df_dummies_accident <- model.matrix(~ accident - 1, data = used_cars_data)  # Dummy variables for 'ext_color'
df_dummies_clean_title <- model.matrix(~ clean_title - 1, data = used_cars_data)  # Dummy variables for 'ext_color'
# Combine the dummy variables with the original numerical features (mileage and model_year)
# df_full <- cbind(df_dummies_brand, df_dummies_model, df_dummies_fuel_type, df_dummies_engine, df_dummies_transmission, df_dummies_ext_color, df_dummies_int_color, df_dummies_accident, df_dummies_clean_title, used_cars_data[, c('mileage', 'model_year','price')])
df_full <- cbind(df_dummies_model, df_dummies_fuel_type, df_dummies_engine, df_dummies_transmission, df_dummies_ext_color, df_dummies_int_color, df_dummies_accident,  used_cars_data[, c('mileage', 'model_year','price')])
```

```{r}
# Check for any NA values in the data
sum(is.na(df_full))  # Returns the number of NA values
sum(is.nan(df_full)) # Returns the number of NaN values
sum(is.infinite(df_full)) # Returns the number of infinite values
```

```{r}
# Perform PCA on the combined dummy variables and numerical features
pca_result <- prcomp(df_full, center = TRUE, scale. = TRUE)

# Calculate explained variance and determine number of components explaining 90% of variance
explained_variance <- pca_result$sdev^2 / sum(pca_result$sdev^2)
cumulative_variance <- cumsum(explained_variance)

# Set the threshold for the variance to keep, e.g., 90%
threshold <- 0.90
n_components <- which(cumulative_variance >= threshold)[1]

# Select the first n_components from the PCA result
selected_components <- pca_result$x[, 1:n_components]

# Combine the selected components with the target variable 'sales'
df_pca <- data.frame(selected_components, price = used_cars_data$price)

# Fit a linear regression model using the selected components
lm_pca <- lm(price ~ ., data = df_pca)

# Summarize the model
summary(lm_pca)
```

```{r}
# Perform both-stepwise regression using p-values
stepwise_model <- ols_step_both_p(lm_pca, penter = 0.05, premove = 0.10)

# View the results of the stepwise selection
summary(stepwise_model)
```



```{r}
missing_fuel_type_rows <- which(is.na(used_cars_data$fuel_type))
# View the rows with missing 'fuel_type' values
used_cars_data[missing_fuel_type_rows, ]
```

```{r}
nrow(df_dummies_brand)
nrow(df_dummies_model)
nrow(df_dummies_fuel_type)
nrow(df_dummies_engine)
nrow(df_dummies_transmission)
nrow(df_dummies_ext_color)
nrow(df_dummies_int_color)
nrow(df_dummies_accident)
nrow(df_dummies_clean_title)
nrow(used_cars_data)
```


```{r}
firstordermodel <- lm(price ~ model_year + mileage + model + 
                      fuel_type + engine + transmission + 
                      ext_col + int_col + accident, 
                      data = used_cars_data)
```


```{r}
vif(firstordermodel)
```


```{r}
# Function to diagnose and handle aliased coefficients
diagnose_multicollinearity <- function(model_data) {
  # Convert categorical variables to factors if not already done
  categorical_vars <- c('brand', 'model', 'fuel_type','engine', 'transmission', 'ext_col', 'int_col','accident','clean_title')
  for (var in categorical_vars) {
    model_data[[var]] <- as.factor(model_data[[var]])
  }
  
  # Print unique levels of categorical variables
  cat("Categorical Variable Levels:\n")
  for (var in categorical_vars) {
    cat(var, ":", length(levels(model_data[[var]])), "levels\n")
  }
  
  # Identify potential source of aliased coefficients
  # Full model with all predictors
  full_formula <- as.formula(paste("price ~ ", 
                                   paste(names(model_data)[names(model_data) != "price"], 
                                         collapse = " + ")))
  
  # Fit the model
  full_model <- lm(full_formula, data = model_data)
  
  # Check for aliased coefficients
  aliases <- alias(full_model)
  
  if (!is.null(aliases$Complete)) {
    cat("\nAliased Coefficients Detected:\n")
    print(aliases$Complete)
    
    # Suggestions for resolving aliased coefficients
    cat("\nPossible Causes and Solutions:\n")
    cat("1. Redundant categorical variables\n")
    cat("2. Perfect linear dependency in predictors\n")
    cat("3. Insufficient variation in some predictors\n")
  }
  
  # Alternative VIF calculation method
  tryCatch({
    # Remove problematic predictors or reduce levels
    reduced_formula <- update(full_formula, . ~ . - model)
    reduced_model <- lm(reduced_formula, data = model_data)
    
    vif_results <- vif(reduced_model)
    cat("\nVariance Inflation Factors (after removal):\n")
    print(vif_results)
  }, error = function(e) {
    cat("Error in VIF calculation:", e$message, "\n")
  })
  
  return(NULL)
}

# Apply the diagnostic function
diagnose_multicollinearity(used_cars_data)
```


```{r}
library(mctest) #for VIF
```

```{r}
imcdiag(firstordermodel, method="VIF")
```



```{r}
library(mctest) #for VIF
firstordermodel=lm(price~model_year+mileage+brand+model+fuel_type+engine+transmission+ext_col+int_col+accident+clean_title, data=used_cars_data)
imcdiag(firstordermodel, method="VIF")
```


# CONCLUSION AND DISCUSSION
## Approach
*Overall, is the approach we took promising? Please elaborate. What different approach or variant of this approach is better?*

## Future work
*What should follow-up work be done next?*

# 5	REFERENCES 
*List all references (e.g., books, journal, conferences, reports, magazines) you are citing in this proposal. Format:*

1.	Data set. **Author(s)**: Taeef Najib, **Title**: Used Car Price Prediction Dataset, **Venue**: Kaggle, **Pages**: N/A, **Month/Year**: May 2022, **Publisher**: Kaggle, **URL**: https://www.kaggle.com/datasets/taeefnajib/used-car-price-prediction-dataset/data
2.	**Author(s)**: Sumeyra Muti , Kazım Yıldız, **Title**: Using Linear Regression For Used Car Price Prediction, **Venue**:International Journal of Computational and Experimental Science and Engineering (IJCESEN), **Pages**:137-141, **Month/Year**: 03/23, **Publisher**:Iğdır University, **URL**:https://doi.org/10.22399/ijcesen.1070505 . 
3.	(Your text here) Author(s), Title, Venue, Pages, Month/Year, Publisher.
4.	(Your text here) Author(s), Title, Venue, Pages, Month/Year, Publisher.




# APPENDIX
Optional part of the proposal. Include any support material you may have, e.g., figures, tables, samples of your data set(s). Make sure to cite any of the Appendix material in the main proposal text. 